Parity games in pushdown systems with parameters

### 2023-06-12 reachability/article.tex
up

\title{Parity games in pushdown systems with pebbles}

\titlerunning{Parity games in pushdown systems with pebbles}

@@

\begin{abstract}
	We investigate the decidability and complexity of
	parity games over an extension of pushdown systems.
	More specifically, we consider pebble pushdown systems, in
	which pebbles can be used to remember some stack content already visited,
	and compared against.
	We determine decidability of the problem, and provide a nonelementary lower bound. 
\end{abstract}

@@

In Section 2, we provide general notations and preliminary definitions. Section 3 will deal with the introduction of 
parametric pushdown games.
% Section 4 is devoted to using pebble alternating tree automata with parity acceptance condition to solve parametric pushdown system parity games. We leave the proof of decidability itself  for Section 5.
Section 6 at last shows a nonelementary lower bound for the problem.

@@

% \paragraph{Reachability Games}
\paragraph{Reachability Games}

@@

% \paragraph{Parity Games}
\paragraph{Parity Games}

@@

\section{Pebble pushdown games}

In this section, we extend pushdown systems with a set of pebbles. 
Pebble pushdown systems will have the ability to lift or drop pebbles on their universe of stack contents according to a strict stack discipline. A \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available) and a \textit{lift} pops the last dropped pebble. One can think of a pebble as a register that can store a stack content.

\begin{definition}
An $n$-pebble pushdown system ($n$-PPDS) is a tuple $\mathcal{Z} = (Q, \Gamma,  \Delta )$
where:
\begin{itemize}
\item $Q$ is a {\em finite set of control states},
\item $\Gamma$ is a {\em finite stack alphabet},
% \item $\$ \notin W$ is an {\em initial stack symbol},
\item  $\Delta   \subseteq  Q  \times (\Gamma \biguplus \{ \$ \})  \times \{ 0, 1, \ldots, n\} \times \mathcal{P}(\{ 1, \ldots, n\}) \times Q  \times (\Gamma^* \cup \{ \! \! \text{\textit{lift}}, \text{\textit{drop}}\})$ is a {\em finite transition relation}, where ${\$ \notin \Gamma}$ is some bottom of stack symbol.
\end{itemize}
\end{definition}

A transition in an $n$-pebble pushdown system works like a transition in a pushdown system that additionally manages the $n$ pebbles.
The idea of a transtion $(q, a, i, S, q', \theta) \in \Delta$
is that, if the automaton $\mathcal{Z}$ is in state $q$ with pebbles $1,\ldots, i$ dropped \---- or without pebble dropped if $i = 0$ \---- with top stack symbol $a$, and stack content which corresponds to the stack contents of the pebbles from $S$ and only these pebbles, then
$\mathcal{Z}$ goes to state 
$q'$ and makes modifications to the stack or the pebbles according to
$\theta$.

A {\em pebble set} of $\mathcal{Z}$ is a set $P \subseteq \{ 1, \ldots, n\}$. For a stack alphabet $\Gamma$, a 
{\em $P$-pebble assignment} is a function which maps each $j \in P$ to a word in $\$\Gamma^*$.
For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, w, f )$, where 
$q$ is a state,
$w \in \$\Gamma^*$, and
$f$ a $\{ 1, \ldots , i \}$-pebble assignment \---- $f$ is the $\emptyset$-pebble assignment in case $i=0$. 
We call  $q$ the current state, $w$ the current stack, and
$f$ the current pebble assignment. 
We also write $(q,w, w_{1} , \ldots , w_{i} )$ if 
$f(j) = w_j$, for each
$j \leq i$.

The set of configurations is the union of all $i$-configurations for $0 \leq i \leq n$. The transition graph of $\mathcal{Z}$ is $(V,E)$, where $V$ is the set of configurations. %and $E$ is such that
Moreover, 
for all $(q, a, i, S, q', \theta) \in \Delta$, with $a \in \Gamma$ (respectively $ \! a = \$ $),
 for all words $w \in \$ \Gamma^*$, and for all $\{ 1, \ldots , i \}$-pebble assignments

 \begin{samepage}
\noindent
 $f$ such that $f(j) = wa$ (respectively $f(j)= \$ $) for each
$j \in S$ 	and
 $f(j) \neq wa$ (resp. $f(j) \neq \$ $) for each $j \in \{ 1, \ldots , i \} \setminus S$,
%the following holds
we include in $E$ the following transitions:
\begin{itemize}
\item if $\theta \in \Gamma^*$, 
$((q,wa,f), (q',w\theta, f)) \in E $ \\
(respectively, $((q,\$,f), (q',\$\theta, f)) \in E $),
\item if $ \theta = \text{\textit{drop}}$, 
$((q,wa,f), (q',wa, f')) \in E $  \\
(respectively, $((q,\$,f), (q',\$, f')) \in E $),\\
where $f'$ is the $\{ 1, \ldots , i, i+1 \}$-pebble assignment such that
$f'(j) = f(j)$ for each
$j \leq i$, and $f'(i+1) = wa$ (respectively, $f'(i+1) = \$$), and
\item if $\theta = \text{\textit{lift}}$, and $ i \in S $%$i>0$
, 
$((q,wa,f), (q',wa, f')) \in E $ \\ 
(respectively, $((q,\$,f), (q',\$, f')) \in E $),\\
where $f'$ is the $\{ 1, \ldots , i - 1 \}$-pebble assignment such that
$f'(j) = f(j)$ for each
$j < i$.
\end{itemize}
\end{samepage}

This defines a graph  $(V,E) $. 
%
To obtain
a parity game, it remains to define the sets $V_0$ and $V_1$ associating the vertices
to the two players, and the priorities of the configurations. One fixes a disjoint
union
 $Q = Q_0  \biguplus Q_1 $. Then 
 $V_0$ is the set of configurations with current state in $Q_0$ and 
 $V_1$ is the set of configurations with current state in $Q_1$.
The mapping  $\Omega$ is first
defined on $Q$, then  $\Omega (q,w,f) =  \Omega (q)$,  for all words $ w  \in \$ \Gamma^*$, all pebble assignments $f$, and 
$q  \in  Q$. So the player and the
priority only depend on the control states of $\mathcal{Z}$. 
This defines the {\em $n$-pebble pushdown game} $ \mathcal{G}_{\mathcal{Z}, Q_0,Q_1,\Omega}$.
The starting configuration is defined if we also fix an initial state 
$q_I  \in  Q$ as $v_I = (q_I, \$, f_I)$, where $f_I$ is the $\emptyset$-pebble assignment. \newline

The main problem we are interested in in this article is the following decision problem.

\problemx{Solving pebble pushdown parity game}
{A pebble pushdown system $\mathcal{Z}= (Q,\Gamma, \Delta )$, a disjoint
union $Q = Q_0  \biguplus Q_1 $, a priority mapping $\Omega$, an initial state $q_I \in Q$% and initial stack symbol $w_I \in W$
.}
{Does player $0$ have a winning strategy from $(q_I,\$, f_I)$ in the  parity game
$ \mathcal{G}_{\mathcal{Z}, Q_0,Q_1,\Omega}$ ?}
%$(\mathcal{G}_{\mathcal{Z}}, \Omega)$ ?\newline}

\section{Tree Automata}

In this section, we use pebble alternating two-way tree automata to solve pebble pushdown parity games. 
We introduce first alternating two-way tree automata, then pebble alternating two-way tree automata,
then explain the reduction. Decidability questions concerning pebble alternating two-way tree automata
are dealt with in Section~\ref{section decidability}.

\subsection{Alternating two-way tree automata}

Given a finite set $\Gamma$ of directions, a $\Gamma$-tree is a prefix closed set 
$T  \subseteq \Gamma^*$,
i.e., if $x.d  \in  T$, where 
$x  \in  \Gamma^*$ and $d  \in  \Gamma $, then $x  \in  T $. 
We will sometimes forget the “.” of the concatenation. 
The elements of $T$ are called {\em nodes}, the
 empty word  $\epsilon$
% word $\$ $
  is the {\em root} of $T $. For every $x.d  \in  T$, $d  \in  \Gamma$ the node $x$ is the
unique {\em parent} of $x.d$, and $x.d$ is a {\em $d$-child} of $x$. The {\em direction} of a node $x.d$
$( \neq   \epsilon )$ 
%$(\neq \$ )$
is $d$. 
The full infinite tree is 
$T = \Gamma^*$. A {\em path (branch)} of a tree $T$ is
a sequence  $\beta   \in  T^{\infty}$  such that 
$ \beta  = u_0 u_1 \ldots u_n$ or 
$ \beta  = u_0 u_1 u_2 \ldots$, where
$u_0 =  \epsilon$  and
 $\forall i < n,  \exists d  \in  \Gamma$, $u_{i+1} = u_i.d$. 
 A path can be finite or infinite.

For a finite set $X$, $\mathcal{B}^+ (X)$ is the set of positive Boolean formulas
over $X$ (i.e., Boolean formulas built from elements in $X$ using only $\wedge$ and $\vee$),
where we also allow the formulas $\text{\textit{true}}$ and $\text{\textit{false}}$. For a set $Y  \subseteq  X$ and a
formula $ \vartheta   \in  \mathcal{B}^+ (X)$, we say that $Y$ satisfies $ \vartheta$  iff assigning $\text{\textit{true}}$ to elements in
$Y$ and $\text{\textit{false}}$ to elements in $X \setminus Y$ makes  $\vartheta $ true.

The set $\text{\textit{types}}(\Gamma) = \{r \} \cup \Gamma$ 
describes the possible types of a node in the full infinite tree $\Gamma^*$. 
Here $r$ stands for the root, $\gamma \in \Gamma$ for a $\gamma$-child.

To navigate through the tree let $ext(\Gamma ) := \Gamma \biguplus \{ \epsilon , \uparrow \}$ be the extended set of directions: The symbol $\uparrow$ means “go to parent node”, 
and  $\epsilon$  means “stay on the present
node”. Formally we define $ \forall u  \in \$ \Gamma^* , d  \in  \Gamma, u.\epsilon  = u$ and $u.d. \! \uparrow= u$, while the node 
$ \epsilon . \! \uparrow $ is not defined.

\begin{definition}
An alternating two-way automaton over
 $\Gamma$-trees (ATWA) is a tuple
$\mathcal{A} = (Q,  \delta , q_I , 
\Omega)$ where
\begin{itemize}
\item $Q$ is a {\em finite set of states},
\item $\delta  : Q  \times  
%
\text{\textit{types}}(\Gamma)
%W \cup \$
  \to  \mathcal{B}^+ (Q \times ext(\Gamma))$ is a {\em transition function},
\item $q_I$ is an {\em initial control state}, and
\item $\Omega: Q \to \{ 0, 1, \ldots, m\}$ is a {\em priority function}.
\end{itemize}
\end{definition}

The idea of a transition $ \delta (q, a) = ( q',\uparrow ) \wedge (q'',d)$ with $a,d\in\Gamma$ is the following: 
if the automaton
$\mathcal{A}$ is in state $q$ on the node $x.a$ of the 
tree $T$, it will send a “copy” of $\mathcal{A}$  in state $q'$ to the node $x$ and
another “copy” in state $q''$ to $x.a.d$. 
After that the two copies are
running independently. They may come again to the same node with two different
states.

More precisely a run of an alternating two-way automaton $\mathcal{A}$ over a tree
is
a tree $T_r$ in which every node is labelled by an
element of $Q \times \Gamma^*$. The latter tree is like the unfolding of the run, its structure
is different from the tree $\Gamma^*$, and its set of directions range rather on the naturals smaller than $|Q|*|ext(\Gamma)|$; a node in $T_r$ labelled by $(q,x)$ describes a “copy”
of the automaton that is in state $q$ and is situated at the node $x$ of $\Gamma^*$. 
Note
that many nodes of $T_r$ can correspond to the same node of $\Gamma^*$, because the
automaton can come back to a previously visited node. 
The label of a node
and its successors have to satisfy the transition function,
that is, for a node $y$ of $T_r$ labelled by $(q,x)$, and $\delta(q,t)= \vartheta$ where $t$ is the type of $x$, there is a (possibly empty) set 
 $Y = \{ (q_0,d_0),(q_1,d_1), \ldots (q_\ell, d_\ell) \}  \subseteq  Q \times ext(\Gamma)$, such that $Y$ satisfies  $\vartheta $, and for all
$(q_i,d_i)  \in  Y$, the node $y$ has a corresponding son $y.i$ labelled by $(q_i,x.d_i)$. Finally, the root of $T_r$ is labelled by $(q_I,\epsilon)$.

Remember that $x.d$ can be $x. \epsilon$  or $x. \! \uparrow$, and the latter is defined only if 
%$x \neq \$ $.
$x  \neq   \epsilon $.
So the run cannot go up from the root of the input tree. Note that it cannot use a
transition  $\delta (q, a) = \text{\textit{false}}$ since the formula $\text{\textit{false}}$ cannot be satisfied.
A run $T_r$ is accepting if all its infinite paths satisfy the parity acceptance 
condition
given by the priority function. 
% The finite paths of a run that end with a transition $ \vartheta $ equivalent to  $true$, which is viewed as successful termination. 
An infinite path $ \beta   \in  T_r^{\omega}$  satisfies the acceptance condition iff the smallest priority appearing infinitely often in
%this path
the sequence $\gamma$ of corresponding labels
 is even, i.e.
%$\min ~ \text{\textit{Inf}}( \Omega (r( \beta ))) $
 $\min(\{ \Omega(q) \mid (q,x) \in \text{\textit{Inf}}(\gamma) \} $ is even. 
Such
a path in the run consists of following only one “copy” of the automaton.

% The generalization of twa with nested pebbles came recently, by two independent motivations: On the one hand, with the advancement of XML theory, finite state recognizers (with name n-pebble tree automata) were used in [21] to show that the XML typechecking problem is decidable. On the other hand, the concept of n-pebble tree-walking automata (n-ptwa) were defined in [9] to recognize first-orderlogic on trees.  Later, in [10]n-ptwa were extended with a more general pebble handling.

\subsection{Alternating two-way tree automata with pebbles}

Informally an $n$-pebble alternating two-way tree automaton works much like an alternating two-way tree automaton, but extended with a fixed set of pebbles, numbered from $1$ to $n$ that it can place on the tree. 
% At each time, pebbles $1,\ldots, i$ are placed on some nodes of the tree, for some $i \in \{1, 2, \ldots , n \}$, or no pebble are placed on the tree. 
To navigate through the tree the automaton can stay at the current node, move to its parent, to its 
 children, it can lift a pebble or place a pebble on the current node.
 Which of these transitions can be applied depends on the current state, the set
 of pebbles at the current node, 
and the type of the current node (root, $\gamma$-child for $\gamma \in \Gamma$). 
Operations $\text{\textit{lift}}$ and $\text{\textit{drop}}$ manage the
pebbles much like in pebble pushdown systems.
We indicate the possible kinds of moves of a pebble automaton by elements of the set 
$ext_2(\Gamma) = ext(\Gamma) \biguplus \{ \text{\textit{lift}}, \text{\textit{drop}}\}$.
% Clearly $\text{\textit{drop}}$ refers to dropping a pebble and $\text{\textit{lift}}$ to lifting a pebble. Note that the notation is  unambiguous by virtue of the stack discipline. 
% Finally recall that if $S$ is a set then $\mathcal{P}(S)$ denotes the powerset of $S$.

\begin{samepage}
\begin{definition}
For $n\geq 0$, an $n$-pebble alternating two-way automaton ($n$-PATWA) over  
$\Gamma$-trees is a tuple
$\mathcal{A} = (Q,   
\delta , q_I , 
\Omega)$ where
\begin{itemize}
\item $Q$ is a {\em finite set of states},
\item $\delta  : Q  \times \{ 0, 1, \ldots, n\} \times \mathcal{P}(\{ 1, \ldots, n\}) \times 
\text{\textit{types}}(\Gamma)
  \to  \mathcal{B}^+ (Q \times ext_2(\Gamma))$ is a {\em transition function},
\item $q_I$ is an {\em initial control state}, and
\item $\Omega : Q \to \{ 0, 1, \ldots, m\}$ is a {\em priority function}.
\end{itemize}
\end{definition}
\end{samepage}

The idea of a transition in a $n$-pebble alternating two-way tree automaton works like a transition in an alternating two-way tree automaton,
except the transition has additional conditions on the pebble assignment, and additional moves for taking care of the pebble.
Consider for instance 
that the automaton
 $\mathcal{A}$ is on the node $x.a$, in state $q$, pebbles $1, 2$ are on the tree, and exactly pebble $1$ is placed on the node $x.a$.
Then a transition
$\delta(q, 2, \{ 1\}, a) =  (q',\uparrow) \wedge (q'',\text{\textit{drop}})$
 will send a “copy” of $\mathcal{A}$  in state $q'$ to the node $x$ and
another “copy” in state $q''$ to $x.a$ with both pebble $1$ and pebble $3$ placed on the node. 
After that the two copies are
running independently.

%A {\em pebble set} of $\mathcal{A}$ is a set $P \subseteq \{ 1, \ldots, n\}$. 
For a tree structure $T$, a 
{\em $P$-pebble assignment} for a pebble set $P$ is a function which maps each $j \in P$ to a node in $T$.
For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, x, f )$, where $x$ is a node, $q$ a state and
$f$ a $\{ 1, \ldots , i \}$-pebble assignment \---- $f$ is the $\emptyset$-pebble assignment in case $i=0$. 
We call $q$ the current state, $x$ the current node, and
$f$ the current pebble assignment. 
We also write $(q, x, x_{1} , \ldots , x_{i} )$ if 
$f(j) = x_j$, for each
$j \leq i$.

Then, formally, a run of an $n$-pebble alternating two-way automaton $\mathcal{A}$
over a tree is a tree $T_r$ in which every node is labelled by 
an $i$-configuration,
for some $i \in \{0, 1, \ldots, n\}$. A node in $T_r$ labelled by $(q, x, x_{1}, x_{2}, \ldots, x_{i})$ describes a “copy” of the automaton that is in state $q$, and is situated on the node $x$ of $\Gamma^*$, with the pebbles up to $i$ already dropped, with the $j$-th pebble for $j \leq i$ dropped on the node $x_{j}$.

The label of a
node and its successors have to satisfy the transition function. % However, recall there are restrictions on the drop-operation and the lift-operation : only
% the pebble with the number $i$ can be lifted and only the pebble with number $i - 1$ can be
% placed.
Formally, this works much the same way as for alternating two-way tree automata; 
that is, the root of $T_r$ is
labelled by $(q_I,\epsilon, f_I )$, and, for a node $y$ of $T_r$ labelled by an $i$-configuration $(q, x, f)$, and 
for
$\delta(q, i, S, t) = \vartheta$ where $t$ is
the type of $x$, 
and $f(j) = x$ for each $j \in S$, and $f(j) \neq x$ for each $j \in \{1,\ldots,i \} \setminus S$,
there is a (possibly empty) set 
$Y \subseteq Q \times ext_2(\Gamma )$, such that $Y$ satisfies $\vartheta$,
and for all $(q',d) \in Y$, the node $y$ has a corresponding son
labelled with %$(x.d, q', f')$, where
\begin{itemize}
\item if $d \in ext(\Gamma)$,
the $i$-configuration $(q ', x.d , f)$,
%
\item if $d = \text{\textit{drop}}$, the $(i+1)$-configuration  $ (q', x, f')$, where the domain of $f'$ is $\{ 1, \ldots, i, i+1\}$, $f'$ is equal to $f$ on the domain of $f$, and $f'(i+1) =x$, and
%
\item if $d = \text{\textit{lift}}$
and $f(i) = x$, i.e. if the pebble $i$ is on the current node%
, the $(i-1)$-configuration $(q', x, f')$, where $f'$ is the restriction of $f$ to $\{1, \ldots, i-1 \}$.
\end{itemize}

A run $T_r$ is {\em accepting} if all its infinite paths satisfy the parity acceptance 
condition 
given by the priority function $\Omega$.
% Again finite paths end with a transition with formula $ \vartheta $ equivalent to $\text{\textit{true}}$, which is viewed as successful termination. 
An infinite path $ \beta   \in  T_r^{\omega}$  satisfies the acceptance condition iff the smallest priority appearing infinitely often in
%this path
the sequence $\gamma$ of corresponding labels
 is even, i.e.
%$\min ~ \text{\textit{Inf}}( \Omega (r( \beta ))) $
 $\min(\{ \Omega(q) \mid (q,x,f) \in \text{\textit{Inf}}(\gamma) \} $ is even.

By a pebble alternating tree-walking automaton (PATWA) we mean an $n$-PATWA for some $n$. 
%Note that a $n$-pebble automaton without pebbles is just an alternating tree-walking automaton.

\subsection{Reduction details}

Alternating two-way tree automata with parity winning conditions have already been used to solve parity games on pushdown graphs in \cite{cachat2002two}, with a technique first developed to solve the model-checking problem in \cite{kupferman2000automata}. The main
idea was to reduce a model-checking problem to an emptiness problem for the class of alternating two-way parity tree automata, and was then adapted to work on pushdown parity games.
%
%
This leads to the following reduction: given a 
parity game on pushdown graph $\mathcal{G}$ and a starting configuration $v_I$, we can
effectively construct an ATWA $\mathcal{A}$
such that
player $0$ has a winning strategy from $v_I$ in $\mathcal{G}$ iff $\mathcal{A}$ accepts the full
infinite tree $\Gamma^*$. The main idea of our reduction is that, when pushdown systems and ATWA are extended with a set of pebbles to manage, the technique used %is conserved.
still works.
We refer the reader to Appendix~\ref{appendix tree reduction} for more details on the proof.

>>>>

\title{Parity games in pushdown systems with parameters}

\titlerunning{Parity games in pushdown systems with parameters}

@@

\begin{abstract}
	We investigate the decidability and complexity of
	parity games over an extension of pushdown systems.
	More specifically, we consider parameters pushdown systems, in
	which parameters can be instantiated to some word over a stack alphabet,
	and compared against the current stack content.
	We determine decidability of the problem, and provide a nonelementary lower bound. 
\end{abstract}

@@

In Section 2, we provide general notations and preliminary definitions. 
Section 3 will deal with the introduction of  parametric pushdown games.
% Section 4 is devoted to using pebble alternating tree automata with parity acceptance condition to solve parametric pushdown system parity games. We leave the proof of decidability itself  for Section 5.
Section 4 is Reduction to higher-order pushdown automata
Section 5 at last shows a nonelementary lower bound for the problem.

@@

% \paragraph{Reachability Games}
\subsection{Reachability Games}

@@

% \paragraph{Parity Games}
\subsection{Parity Games}

@@

\section{Parametric pushdown games}

\textcolor{black}{Parametric pushdown automata extend pushdown automata by allowing the stack to be compared against parameters that can be assigned values which are words over the stack alphabet.}
A parametric pushdown automaton is then a finite automaton extended with a finite set of parameters $P$  and with a stack that can be manipulated by pushing or popping stack symbols and such that, moreover, the automaton can use the top of the stack, or check that the stack content corresponds to a parameter,  to decide which transition to take next.\\

\par\noindent\ignorespacesafterend
Formally, a {\em parametric pushdown automaton} ({\em PPDA} for short) 
is a tuple 
$\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}, F),$ where
\begin{itemize}
\item $Q$ is a non-empty finite {\em set of  states},
\item $\Gamma$ is a non-empty finite {\em  stack alphabet},
\item $P$ is a finite {\em   set of parameters}, \textcolor{black}{with $\Gamma \cap P = \emptyset$},
% \item  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Gamma^*$ is finite {\em  set of rules},
 \item  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Op(\Gamma)$ is finite {\em  set of rules},
\item $q_{init}\in Q$ is an {\em initial  state}, 
\item $ \gamma_{init} \in \Gamma$ is an {\em initial stack symbol}, and
\item $F\subseteq Q$ is a {\em set of final  states}.
\end{itemize}

\par\noindent\ignorespacesafterend
The {\em size} of $\mathcal{Z}$ is defined as
$|\mathcal{Z}| = |Q| + |\Gamma| + |P| + |R| $.
We also refer to $\mathcal{Z}$ as an $n$-parametric pushdown automaton if $|P| = n$.
A {\em stack content} is a word from $ \Gamma^*$. 
As before we write the top
of the stack at the right of the word%(we are considering suffix rewriting as in Chapter 15)
. 
% A configuration is a  state and a stack content: $(q, w)$,
% which we rather write $q(w)$, where $w  \in  W^*$, $q  \in Q$. 
By $\Conf(\mathcal{Z})=Q\times  \Gamma^*$ we denote the set of
{\em configurations} of $\mathcal{Z}$. As usual we rather write $q(w)$ instead of $(q, w)$.
A {\em parameter valuation} is a function $\mu$ from $P$ to $\Gamma^*$.

\begin{samepage}
%\begin{definition}
A parametric pushdown automaton  $\mathcal{Z}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init},F)$ 
and a parameter valuation $\mu: P\rightarrow \Gamma^*$ induce the 
%labeled 
transition system  
$T_{\mathcal{Z}}^{\mu} = (\Conf(\mathcal{Z}), 
%\Lambda_{\mathcal{Z}}, 
\rightarrow_{\mathcal{Z},\mu})$ where
% $\Lambda_{\mathcal{Z}} = \Gamma \cup P$ and
for all $ q, q'  \in  Q$, 
for all $ w,w' \in \Gamma^* $, and 
for all $ a  \in  \Gamma$, 
\iffalse
\begin{eqnarray*}
q(wa) \xrightarrow{%a
}_{\mathcal{Z},\mu} q'(w') & \text{if} &  (q, a, q' , op)  \in  R, \\
q(\mu(p) ) \xrightarrow{%p
}_{\mathcal{Z},\mu} q'(\mu(p) %\textcolor{red}{\cdot} 
w) & \text{if} &  (q, p , q' , w)  \in  R.
\end{eqnarray*}
% \end{definition}
\fi
$q(wa) \xrightarrow{%a
}_{\mathcal{Z},\mu} q'(w')$
if there exists a rule in $R$
 of the form $(q,a,q', op)$ or $(q,p,q', op)$ with $\mu(p)=wa$,
such that
		% $\text{\textit{top}}(w) = a$
		% and 
% $w' = \theta(w)$.
either of the following holds
\begin{itemize}
\item $ op = \text{\textit{push}}^\gamma$ and $w' = wa\gamma$,

\item $ op = \text{\textit{pop}}$ and $w' = w$, or

\item $ op = \text{\textit{skip}}$ and $w' = wa$.
\end{itemize}
\end{samepage}

A {\em $\mu$-run} from $q_0(a_0)$ to $q_n(a_n)$ in $\mathcal{Z}$ is a corresponding
path in the transition system 
$T_{\mathcal{Z}}^\mu$ 
induced by $\mathcal{Z}$ and $\mu$.
%
%and will be noted
%$q_0(z_0) \xrightarrow{\lambda_0}_{\mathcal{Z},\mu} q_1(z_1) \xrightarrow{\lambda_1}_{\mathcal{Z},\mu} \cdots \xrightarrow{\lambda_{n-1}}_{\mathcal{Z},\mu} q_n(z_n)$.
%
As with PDA, we say $\pi$ is {\em accepting} if $q_0 = q_{init}$, $a_0=\gamma_{init}$,
and $q_n\in F$.
%
% We say {\em reachability holds} for $\mathcal{Z}$ if there is a $\mu$ such that there is a $\mu$-run in $\mathcal{Z}$ from $q_{init}(\gamma_{init})$   to some configuration $q(w)$ for some $q\in F$ and some $w \in \Gamma^*$.
% We refer to Figure~\ref{example_pda} for an instance of a PDA for which reachability holds.

In the particular case where $P=\{p\}$ is a singleton for some parameter $p$ and $\mu(p)=u \in \Gamma^*$,
we prefer to write $q(w)\xrightarrow{%\lambda
}_{\mathcal{Z}, u}q'(w')$
to denote $q(w)\xrightarrow{%\lambda
}_{\mathcal{Z}, \mu}q'(w')$ and
will call the $\mu$-run an {\em $u$-run}. In case the automaton $\mathcal{Z}$ is obvious from context, we write $\xrightarrow{%\lambda
}_{\mu}$ (resp. $\xrightarrow{%\lambda
}_{u}$)
instead of $\xrightarrow{%\lambda
}_{\mathcal{Z}, \mu}$ (resp. $\xrightarrow{%\lambda
}_{\mathcal{Z},u}$). \\

\subsection{PPDA reachability}

Note that the reachability problem for parametric pushdown automata consists in
the following decision problem.

\problemx{$n$-PPDA reachability}
{An $n$-PPDA $\mathcal{Z}$.}
{Does 
%reachability hold for
an accepting $\mu$-run for some $\mu\in(\Gamma^*)^P$ exist in
 $\mathcal{Z}$?\newline}

% \subsubsection{Parametric pshdown games}
\subsection{PPDA games}

We are moreover interested in games played on parametric pushdown automata's  
 transition systems.
Recall that given a 
%labeled 
transition system,
one needs only to provide %a mapping $\Omega$ and 
a partition of the set of configuration $S$ into two sets $S_0$ and $S_1$ to obtain an arena. \\

Thus, for an  $n$-parametric pushdown automaton 
$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init},F)$,
given a partition of
$Q$ into $Q_0$ and $Q_1$, 
we naturally partition the configurations of
%the transition system of 
 $\mathcal{Z}$
% $T^\mu_{\mathcal{Z}}$
into
$\Conf_{\mathcal{Z},0}=Q_0\times \Gamma^*$
and
$\Conf_{\mathcal{Z},1}=Q_1\times \Gamma^*$.

With these notations in mind one can define the arena
$$
A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
=
(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$$
induced by a PPDA $\mathcal{Z}$, a partition of its set of states, and a
parameter valuation $\mu: P\rightarrow \Gamma^*$.

Given additionally a partition of
$P$ into $P_0$ and $P_1$, 
we define 
$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
as
the arena that contains both the configurations and transitions of
the parameter valuation arena
$A_{P_0,P_1,\Gamma^*}$, as see on page~\pageref{parametric arena},
and that moreover contains the configurations and transitions of
$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$ 
\---- albeit with configurations additionally including $\mu$ in the tuple to avoid confusion \----
for
all parameter valuations $\mu: P\rightarrow \Gamma^*$.
Moreover, for every configuration $\mu$ in $A_{P_0,P_1,\Gamma^*}$,
we add a transition
from $\mu$ towards 
$q_{init}(\gamma_{init}, \mu)$ in $A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$. \label{PPDA reachability game}

Given a 
priority function $\Omega: Q \to [0, m]$,
we consider the function
$\widehat{\Omega}: \Conf(\mathcal{Z}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$
where we set
$\widehat{\Omega}(q, w, \mu) = \Omega(q)$
for all
$w \in \Gamma^*$
and
for all $\mu \in (\Gamma^*)^{P}$,
and
$\widehat{\Omega}(\mu) = 0$
for all $\mu \in (\Gamma^* \cup \bot)^{P}$.

% $T_{\mathcal{Z}} = (\Conf(\mathcal{Z}), \rightarrow_{\mathcal{Z},\mu})$

\iffalse
\textcolor{red}{
Reachability %, B\"uchi
 and parity games on $\mathcal{G}_{(\mathcal{Z}, \mu,P_0, P_1, Q_0 , Q_1, \Omega)}$
are then clearly defined. They have the flaw however of not taking into account the
way the values assigned to the parameters is chosen. How values are assigned to
the parameters is the subject of the next Section.
}
\fi

We are interested in the following games and problems.

\problemx{Parametric pushdown reachability game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
, and $ P = P_0 \biguplus P_1$.}
{Does player $0$ have a winning strategy from $\mu_\bot$ in the reachability game \\
$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{F \times \Gamma^* \times (\Gamma^*)^P})$ ?\newline}

\problemx{Parametric pushdown parity game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
, and $ P = P_0 \biguplus P_1$, 
and a priority function $\Omega: Q \to [0, m]$%
%which naturally lead to a mapping on $\Conf(\mathcal{Z})$
%, and an initial state $q_{init} \in Q$% and initial stack symbol $w_{init} \in W$
.}
{Does player $0$ have a winning strategy from $\mu_\bot$ in the parity game \\
$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$ ?\newline}

\iffalse
\problemx{Parametric pushdown reachability game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, a disjoint
union $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
.}
{Does player $0$ have a 
strategy $\sigma_0$ for the parameter valuation game  such that
for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
player $0$ has a
winning strategy from $q_{init}(\gamma_{init})$ for the reachability condition on
$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$ ?\newline}

\problemx{Parametric pushdown parity game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, a disjoint
union $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
, 
a priority mapping $\Omega: Q \to \{0, \ldots m\}$%
%which naturally lead to a mapping on $\Conf(\mathcal{Z})$
%, and an initial state $q_{init} \in Q$% and initial stack symbol $w_{init} \in W$
.}
{Does player $0$ have a 
strategy $\sigma_0$ for the parameter valuation game  such that
for all player $1$ strategies $\sigma_1$ for the parameter valuation game, 
player $0$ has a
winning strategy from $q_{init}(\gamma_{init})$ for the parity condition on
$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$ ?\newline}
\fi

We write {\sc $n$-parametric reachability parity game}
and
{\sc $n$-parametric pushdown parity game} if 
 the number of parameters $n=|P|$ of $\mathcal{Z}$ is fixed by the problem.
